---
title: "Homework 5"
author: Daniel Dulaney
date: 
output: distill::distill_article
editor_options: 
  chunk_output_type: console
---

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(ggeasy)
```


## 1.

```{r}
# read data in and convert 3 columns to type factor
crabs <- 
  read.table(here("hw05_summarizing-data", "crabs.txt"),
           header = TRUE) %>% 
  as_tibble() %>% 
  mutate_at(vars(color, spine, y), as.factor)

# relabel the 3 factors
levels(crabs$color) <- c("light", "medium", "dark", "darker")
levels(crabs$spine) <- c("Both Good", "One Worn/Broken", "Both Worn/Broken")
levels(crabs$y) <- c("No Sattelite", "At least 1 Sattelite")

print(crabs)
```


## 2.

```{r}
table(crabs$satell, crabs$spine)
```

The number in the very top-left tells us that for female horshoe crabs with no
sattelite males, 11 (out of 62) have two good spines. The number in the very
bottom-right tells us that the only crab with 15 sattelites has two broken
or worn spines.


## 3.

```{r}
tab <- table(crabs$color, crabs$spine, crabs$satell)
tab
```

```{r}
# since "darker" is the 3rd level
tab[3, , ]
```

The number in the very top-left tells us that there are no dark-colored crabs
with 0 sattelite males and spines that are both good.


## 4.

```{r}
crabs %>% 
  ggplot(aes(color, fill = as.factor(y))) +
  geom_bar(position = "dodge") +
  coord_flip() +
  ggeasy::easy_remove_legend_title() +
  labs(x = "Female Crab Color")
```

```{r}
crabs %>% 
  ggplot(aes(spine, fill = as.factor(y))) +
  geom_bar(position = "dodge") +
  coord_flip() +
  ggeasy::easy_remove_legend_title() +
  labs(x = "Female Crab Spine Condition")
```

```{r}
crabs %>% 
  ggplot(aes(spine, fill = as.factor(color))) +
  geom_bar(position = "dodge") +
  coord_flip() +
  ggeasy::easy_add_legend_title("Female Crab Color") +
  labs(x = "Female Crab Spine Condition")
```

Male crabs seem to like females who are medium-colored, as well as those with
two good spines.


## 5.

```{r}
crabs %>% 
  mutate(spine = str_c("spine: ", spine),
         spine = fct_relevel(spine, c("spine: Both Good",
                                      "spine: One Worn/Broken",
                                      "spine: Both Worn/Broken"))) %>% 
  ggplot(aes(color, fill = as.factor(y))) +
  geom_bar(position = "dodge") +
  facet_wrap(~spine) +
  ggeasy::easy_rotate_x_labels(angle = 45) +
  ggeasy::easy_remove_legend_title() +
  labs(x = "Female Crab Spine Condition")
```

There are relatively more sattelite-less female crabs among those with one or 
more broken spines.


## 6.

```{r}
crabs %>% 
  group_by(color, y) %>% 
  summarise(Avg = mean(weight),
            Sd = sd(weight),
            Median = median(weight),
            IQR = IQR(weight))
```

The first row gives us the average, standard deviation, median, and 
(3rd quartile - 1st quartile) weight for all light-colored crabs with no sattelite males.


## 7.

```{r}
crabs %>% 
  ggplot(aes(y, weight)) +
  geom_boxplot() +
  geom_point(aes(color = y), position = "jitter") +
  labs(title = "Boxplot for Weight")
```

```{r}
crabs %>% 
  ggplot(aes(y, width)) +
  geom_boxplot() +
  geom_point(aes(color = y), position = "jitter") +
  labs(title = "Boxplot for Width")
```

Female crabs with greater weight and width usually have more sattelites.


## 8.

```{r}
cor(crabs$weight, crabs$width)
```

The correlation between weight and width is r = 0.89

```{r}
crabs %>% 
  ggplot(aes(weight, width)) +
  geom_point(aes(color = color, shape = y)) +
  geom_smooth(method = "lm") +
  labs(title = "Weight vs Width")
```

Weight and width are extremely highly correlated in the positive direction.
